---
title: "Reasearch_Analysis.Rmd"
author: "Tasbiha fatima  Hasan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library( readxl)

```

```{r}
Data_sheet_Project_Hasan <-read_excel("C:/Users/qj6578om/OneDrive - Minnesota State/Desktop/Data Science 211-01/Reseach -Project/Data sheet_Project _Hasan.xlsx")
head(Data_sheet_Project_Hasan)
```


```{r}
water <- Data_sheet_Project_Hasan %>%
  mutate(Intake_L = `Bottle Consumed ( ml)` * 0.5)
head(water)
```




```{r}
water <- Data_sheet_Project_Hasan %>%
  select(-`Reccommneded =2.7 L`) %>%
  mutate(
    Intake_L = `Convert ml *Liters`,
    Status = case_when(
      Intake_L < 2.7 ~ "Low",
      Intake_L == 2.7 ~ "Met",
      Intake_L > 2.7 ~ "High"
    ),
    Difference = Intake_L - 2.7,
    Weekday = weekdays(as.Date(Date)),
    Date = format(as.Date(Date), "%m/%d/%Y")                                         
  ) %>%
  select(Date, Weekday, Intake_L, Status, Difference)                                   
head(water)
```

```{r}
water <- water %>%
  filter(!is.na(Status)) %>%
  mutate(Status = factor(Status, levels = c("Low", "Medium", "High")))
head(water)

```

# this was done to remove N/A 




## Q3 plot the graph 
```{r}
library(tidyverse)

# Sample data (replace with your actual water data)
water <- c(2.0, 1.5, 2.5, 1.5, 2.0, 2.0, 2.0,3.0, 3.0, 2.0, 3.75, 3.0,3.0,3.5 )

# Plot
ggplot(data.frame(water), aes(x = water)) +
  geom_histogram(binwidth = 0.1, fill = "steelblue", color = "black") +
  geom_vline(xintercept = 2.7, color = "red", linetype = "dashed", size = 1) +
  labs(
    title = "Histogram of Daily Water Intake",
    x = "Water Intake (Liters)",
    y = "Number of Days"
  ) +
  theme_minimal()

```
# Q2 Summary Statistics report
```{r}
water<-c(2.0, 1.5, 2.5, 1.5, 2.0, 2.0, 2.0,3.0, 3.0, 2.0, 3.75, 3.0,3.0,3.5 )
mean_intake<-mean(water)
sd_intake<-sd(water)
range_intake<-range(water)
mean_intake
sd_intake
range_intake
```




$p$: proportion of days with water intake of below 2.7 L 
    $H_0:p = 0.5$ vs $H_1:P<0.5$

```{r}
# Number of days under 2.7 liters
x <- 8  

# Total number of days
n <- 14     

# Test if proportion < 0.5 (one-tailed)
prop.test(x, n, p = 0.5, alternative = "greater", correct = FALSE)

```
# Result 

p- value 0.2965
Significance level= 0.05 
Since is greater that 0.05 we fail to reject the the null hypothesis

At the $

$p$-value = 0.2965\alpha = 0.05$, so we reject
$H_0$
At $5\%$ significance level
we have enough data to conclude  that less than of the half of the days my water is below the recommended 2.7 L .This means that my water intake  waS not 
significance low on the majority of days .
## Q4 Reflection 
A major trend I noticed is that my average water intake is below the recommended 2.7 liters per day. Previously, I believed that drinking 8 glasses of water was the standard, but through this project, I learned that 2.7 liters is actually a more evidence-based guideline—exceeding the 8-glass rule. This realization helped me reassess my habits. One challenge is that during winter, Even I feel 
thirsty and I often delay drinking water due to forgetfulness, busyness, or procrastination. Tracking my hydration made me more aware that building consistent habits requires intentional effort and time, especially when daily routines are already established.




To test using t- test : 
```{r}
t.test(
  water,
  mu = 2.7,
  alternative = "lesS"
)

```
$p{-value} = 0.1402 > \alpha = 0.05$. 
Therefore, we fail to reject $H_0$. There is 
not sufficient statistical evidence to conclude that my average daily water 
intake is less than the recommended 2.7 liters per day.

The 95% confidence interval for the true mean ranged from negative infinity to 2.82 L, which includes 2.7 L. This indicates that although my observed intake is lower, there is insufficient statistical evidence to conclude that my true average water intake is less than the recommended amount.

$H_0:\mu\ge 2.7$ vs $H_1:\mu < 2.7$ 

```{r}
library(ggplot2)

#  water data
water <- c(2.0, 1.5, 2.5, 1.5, 2.0, 2.0, 2.0, 3.0, 3.0, 2.0, 3.75, 3.0, 3.0, 3.5)

# Sample mean and CI
n <- length(water)
mean_intake <- mean(water)
sd_intake <- sd(water)
error <- qt(0.975, df = n - 1) * sd_intake / sqrt(n)
lower_ci <- mean_intake - error
upper_ci <- mean_intake + error

# Create data frame for plot
df_plot <- data.frame(
  group = c("My Intake", "Recommended"),
  mean = c(mean_intake, 2.7),
  lower = c(lower_ci, 2.7),   # fixed guideline = no CI range
  upper = c(upper_ci, 2.7)
)

# Plot
ggplot(df_plot, aes(x = group, y = mean)) +
  geom_point(size = 4, color = "blue") +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = 0.15, size = 1) +
  labs(
    title = "Comparison of My Daily Water Intake 
    to the Recommended Level (95% CI) ",
    x = "",
    y = "Water Intake (Liters)"
  ) +
  theme_minimal(base_size = 14)

```

```{r}

t.test(
  water,
  mu = 2.7,
  alternative = "two.sided"
)
```

